Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-12-04 MDPI_submission.tex
Commit 2023-12-04 21:25

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
operating continuously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
2023-04-13. The monitoring site is located near the city centre, and we
expect to be affected by the urban environment. During the study period,
the pyranometer has been independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, when it was verified the
stability of the instrument to within better than \(0.7\%\) relative to
the initial calibration by the manufacturer. Along with SDR, the direct
beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-13. Although,
we have performed a similar analysis to the DNI data, the results are
not presented here, as they lack the appropriate statistical
significance, due to the sorter duration of the data. However, the DNI
data were used as auxiliary data, in the clear sky identification
algorithm (CSid), which is discussed later, for the selection of the
appropriate thresholds. It is noted that despite the capability of the
CSid algorithm to use the DNI as a characterization parameter, we
haven't utilized it here, to avoid any selection bias, due to unequal
length of the two datasets. There are four distinct steps in the
creation of the dataset analysed here: a)~the acquisition of radiation
measurements from the sensors, b)~the data quality check, c)~the
identification of ``clear sky'' conditions from the radiometric data,
and d)~the aggregation of data and trend analysis.

@@

We calculated the linear SDR trends, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR
slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
before the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00022\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
\midrule
All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-7}
0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
\cmidrule{1-7}
0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
-0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & Clear skies & Spring & 82.9\\
\cmidrule{1-7}
\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
\cmidrule{1-7}
0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
\cmidrule{1-7}
0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
\bottomrule
\end{tabu}
\end{table}

>>>>

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
operating continuously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
2023-04-12. The monitoring site is located near the city centre, and we
expect to be affected by the urban environment. During the study period,
the pyranometer has been independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, when it was verified the
stability of the instrument to within better than \(0.7\%\) relative to
the initial calibration by the manufacturer. Along with SDR, the direct
beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-12. Although,
we have performed a similar analysis to the DNI data, the results are
not presented here, as they lack the appropriate statistical
significance, due to the sorter duration of the data. However, the DNI
data were used as auxiliary data, in the clear sky identification
algorithm (CSid), which is discussed later, for the selection of the
appropriate thresholds. It is noted that despite the capability of the
CSid algorithm to use the DNI as a characterization parameter, we
haven't utilized it here, to avoid any selection bias, due to unequal
length of the two datasets. There are four distinct steps in the
creation of the dataset analysed here: a)~the acquisition of radiation
measurements from the sensors, b)~the data quality check, c)~the
identification of ``clear sky'' conditions from the radiometric data,
and d)~the aggregation of data and trend analysis.

@@

We calculated the linear SDR trends, from the departures of the mean
daily values from the daily climatology and for the three sky conditions
(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
only the time series under all-sky conditions; the plots for clear-sky
and cloud-sky conditions, are very similar and are shown in the Appendix
(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
no significant break or change in the variability of the time series.
Other studies for the European region reported a change of the SDR
slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
before the start of our records. It is interesting to note, that for the
observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
thus we can not attribute any major effect on SDR trend to Solar
variability.

@@

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
\midrule
All skies & 0.3756 & 0.000 & 0.0911 & 10250\\
Clear skies & 0.0239 & 0.189 & 0.0172 & 5803\\
Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
\toprule
Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-7}
0.1360 & 2.19e-01 & 0.0515 & 0.2270307 & All skies & Spring & 78.1\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
\cmidrule{1-7}
0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
-0.1450 & 1.49e-01 & 0.0706 & -0.2656359 & Clear skies & Spring & 85.1\\
\cmidrule{1-7}
\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-7}
0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
\cmidrule{1-7}
0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
\cmidrule{1-7}
\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
\cmidrule{1-7}
0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
\bottomrule
\end{tabu}
\end{table}
